---
output: 
  github_document: 
    default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "fig/README-"
)
```

# printy

Over the years, I've written a lot of one-off functions for formatting numbers
in RMarkdown documents. This packages collects them in a single location.

## Installation 📚

You can install printy from github with:

```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("tjmahr/printy")
```

## Formatters ✍ 

`fmt_fix_digits()` prints a number with n digits of precision. R numbers lose
precision when converted to strings. This function converts the numbers to
strings and keeps precision. (It's a wrapper for `sprintf()`.)

```{r fix-digits}
library(dplyr, warn.conflicts = FALSE)
library(printy)
test_cor <- cor(mtcars[, 1:4]) 

# Typical loss of trailing zeroes
test_cor[1:4, 3] %>% round(2) %>% as.character()

test_cor[1:4, 3] %>% fmt_fix_digits(2)
```

`fmt_leading_zero()` removes a leading zero on numbers that are bounded between
−1 and 1, such as correlations or *p*-values.

```{r leading-zero}
fmt_leading_zero(c(-0.3, 0.4, 1))

# Todo: This functions obliterates 0's
fmt_leading_zero(c(0, 0.0))
```

`fmt_minus_signs()` formats negative numbers with a minus sign.

```{r minus-sign}
fmt_minus_sign(c(1, 2, -3, -0.4, -pi))
```

Putting it all together: Print a correlation matrix with 2 digits, no leading
zero and with minus signs.

```{r}
fmt_correlation <- function(xs, digits = 2) {
  xs %>% fmt_fix_digits(digits) %>% fmt_leading_zero() %>% fmt_minus_sign()
}

test_cor %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(".rowname") %>% 
  tibble::as_tibble() %>% 
  mutate_at(vars(-.rowname), fmt_correlation) %>% 
  rename(` ` = .rowname) %>% 
  knitr::kable(align = "lrrrr")
```

### *p*-values 🎣

`fmt_p_value()` formats *p*-values with *n* digits of precision, with no leading
zero, and with very small values being printed with a `<` sign.

```{r}
p <- c(1, 0.1, 0.01, 0.001, 0.0001)
fmt_p_value(p, digits = 2)
fmt_p_value(p, digits = 3)
```

`fmt_p_value_md()` formats *p*-values in markdown with nice defaults. 

* Use 3 digits of precision for values less than .06
* Otherwise, use 2 digits of precision.
* Include *p* in markdown

```{r}
p <- c(1, 0.1, 0.06, 0.059, 0.051, 0.01, 0.001, 0.0001)
fmt_p_value_md(p)
```

These render as: `r paste0(fmt_p_value_md(p), collapse = ", ")`.


### Experimental formatters  🧪

`fmt_effect_md()` is an experimental function for getting model effects
formatted in markdown. You give the function a model, an effect and a string
listing the quantities you want.

```{r}
model <- lm(breaks ~ wool * tension, warpbreaks) 
summary(model)
```


```{r}
# default to: b (beta), e (error), s (statistic), p (p value)
fmt_effect_md(model, "woolB", "besp")
```

`r fmt_effect_md(model, "woolB", "besp")`

```{r}
# Just a subset of them
fmt_effect_md(model, "woolB", terms = "bp")
```

`r fmt_effect_md(model, "woolB", terms = "bp")`

```{r}
# B for labeled b
fmt_effect_md(model, "woolB", terms = "Bp", b_lab = "Wool B")
```

`r  fmt_effect_md(model, "woolB", terms = "Bp", b_lab = "Wool B")`

```{r bi}
# i for interval
fmt_effect_md(model, "woolB", terms = "bi")
```

`r fmt_effect_md(model, "woolB", terms = "bi")`

```{r bSp}
# S for statistic with df
fmt_effect_md(model, "woolB", terms = "bSp")
```

`r fmt_effect_md(model, "woolB", terms = "bSp")`


```{r}
# extra digits (except for p-values; those go through `fmt_p_value_md()`)
fmt_effect_md(model, "woolB", terms = "bep", digits = 6)
```

`r fmt_effect_md(model, "woolB", terms = "bep", digits = 6)`


## Skeletons 🦴

I use `fmt_` for formatting functions. The other convention in the package is
`skel_` to plug values into a formatting skeleton. 

`skel_conf_interval()` creates a confidence interval from two numbers.


```{r}
skel_conf_interval(c(1, 2))
```


`skel_conf_interval_v()` is the vectorized version. It is suitable for working
on columns of numbers.


```{r}
model <- lm(breaks ~ wool * tension, warpbreaks) 

ci_starts <- confint(model)[, 1] %>% 
  fmt_fix_digits(2) %>% 
  fmt_minus_sign()

ci_ends <- confint(model)[, 2] %>% 
  fmt_fix_digits(2) %>% 
  fmt_minus_sign()

skel_conf_interval_v(ci_starts, ci_ends)
```

## Formatting tables from lme4 models 🖇

One thing I've had to do a lot is summarize mixed effects models fit with lme4.
This package provides `pretty_lme4_ranefs()` which creates a dataframe random
effect variances and covariances like those printed by `summary()`.

For example, we can fit the model.

```{r}
library(lme4)
model <- lmer(Reaction ~ Days + (Days | Subject), sleepstudy)
summary(model)
```

`pretty_lme4_ranefs()` creates the following dataframe.

```{r}
pretty_lme4_ranefs(model)
```

Which in markdown renders as

```{r}
knitr::kable(
  pretty_lme4_ranefs(model), 
  align = c("l", "l", "r", "r", "r")
)
```

Here's a dumb model with a lot going on in the random effects.

```{r, warning = FALSE}
model <- lmer(mpg ~ wt * hp + (drat | gear) + (hp * cyl | am), mtcars)
model

knitr::kable(
  pretty_lme4_ranefs(model), 
  align = c("l", "l", "r", "r", "r", "r", "r", "r", "r")
)
```

